CreateFiguresReview

### 2025-07-08 manuscript/Suppl_material/supplementary.Rmd
adjust Suppl. Figure 1

In EHA parlance, the single event of interest is a button press response, *time zero* is defined as target display onset, the  *observation period* is 500 ms long in each trial or repeated measurement, in each trial the participant is *at risk* for response occurrence as long as the response has not occurred yet, the individual starts in an "idle" state in each trial and *transitions* to a "detected" state when a response occurs, all trials have a censoring time of 500 ms, and some trials result in observed event times (those with a RT below 500 ms), while other trials result in response times that are right-censored at 500 ms.

@@

(ref:descr-plot-RT-tasks-caption) RT tasks and EHA concepts. The timer starts at time 0 (black vertical line) and ends at the censoring time (dotted vertical line). In each trial of each condition of each task, an overt response time (RT) is indicated by a circle. (A) Single-button detection data requires a single-event analysis. (B) Two-button discrimination data requires a competing risks analysis. (C) Competing risk data in combination with an stimulus-onset-asynchrony manipulation might require increasing the observation period in each trial. (D) Two-button bistable perception data requires a recurrent event / multistate analysis.

@@

For example, in Supplementary Figure 1B, we present a hypothetical single-subject RT data set from a discrimination experiment with two response buttons and two experimental conditions: A and B. For example, each trial in each condition might start with a fixation screen for 1 s, followed by a blank screen for 1 s, a target display screen for 100 ms in which a female or male face is presented, and finally a blank screen until an overt response is detected or 800 ms have passed. Participants are required to press the left button when identifying a male, and the right button when identifying a female. In Supplementary Figure 1B, correct responses are indicated with black dots, error responses with white dots, and the few right-censored observations with crosses.

@@

Studying the effect of stimulus-onset-asynchrony (SOA) on mean RT is still a popular approach to study the time course of cognitive processes. For example, a target stimulus can be presented with a certain SOA in relation to another stimulus (e.g., prime, mask, distractor, cue). However, mean RT conceals the possibe presence of responses triggered by the non-target stimulus (Panis, 2020; Panis & Schmidt, 2016). For example, in Supplementary Figure 1C a prime is presented before a target with a SOA of -400 ms (upper x-axis labels). To capture responses triggered by the prime (and emitted before target onset) we have to start the timer at trial onset. Also, when RTs are originally recorded relative to target onset, we have to add a constant time to each (here: + 400 ms) in order to include these "negative" RTs in the analysis (lower x-axis labels). In general, all RTs must be positive for EHA to work.

@@

While psychological RT data is typically measured in small, continuous units (e.g., milliseconds), discrete-time EHA treats the RT data as interval-censored data, because it only uses the information that the response occurred sometime in a particular bin of time (x,y]: x < RT $\leq$ y. If we want to use the exact event times, then we treat time as a continuous variable, and let RT be a continous random variable denoting a particular person's response time in a particular experimental condition. Continuous-time EHA does not focus on the cumulative distribution function F(t) = P(RT $\leq$ t) and its derivative, the probability density function f(t) = F(t)', but on the survivor function S(t) = P(RT $>$ t) and the hazard rate function $\lambda$(t) = f(t)/S(t). The hazard rate function gives you the instantaneous *rate* of event occurrence at time point t, given that the event has not occurred yet. Models for continous-time event data come in three forms: parametric models, Cox regression, and piece-wise exponential models [@singerAppliedLongitudinalData2003].

@@

Third, EHA does not discard right-censored observations when estimating hazard functions, that is, trials for which we do not observe a response during the data collection period in a trial so that we only know that the RT must be larger than some value (e.g., the response deadline). This is important because although a few right-censored observations are inevitable in most RT tasks, a lot of right-censored observations are expected in experiments on masking, the attentional blink, and so forth. In other words, by using EHA you can analyze RT data from experiments that typically do not measure response times. As a result, EHA can also deal with long RTs in experiments without a response deadline, which are typically treated as outliers and are discarded before calculating a mean. This orthodox procedure leads to underestimation of the true mean. By introducing a fixed censoring time for all trials at the end of the analysis time window, trials with long RTs are not discarded but contribute to the risk set of each bin.

>>>>

In EHA parlance, the single event of interest is a button press response, *time zero* is defined as target display onset, the  *observation or follow-up period* is 500 ms long in each trial or repeated measurement, in each trial the participant is *at risk* for response occurrence as long as the response has not occurred yet, the individual starts in an "idle" state in each trial and *transitions* to a "detected" state when a response occurs, all trials have a censoring time of 500 ms, and some trials result in observed event times (those with a RT below 500 ms), while other trials result in response times that are right-censored at 500 ms.

@@

(ref:descr-plot-RT-tasks-caption) RT tasks and EHA concepts. The timer starts at time 0 (black vertical line) and ends at the censoring time (dotted vertical line). In each trial of each condition of each task, an overt response time (RT) is indicated by a circle. (A) Single-button detection data requires a single-event analysis. (B) Two-button discrimination data requires a competing risks analysis. (C) Competing risk data in combination with an stimulus-onset-asynchrony manipulation might require increasing the observation period in each trial. (D) Two-button bistable perception data requires a recurrent event / multistate analysis. Black bars indicate stimlus presentation.

@@

In Supplementary Figure 1B, we present a hypothetical single-subject RT data set from a discrimination experiment with two response buttons and two experimental conditions: A and B. For example, each trial in each condition might start with a fixation screen for 1 s, followed by a blank screen for 1 s, a target display screen for 100 ms in which a female or male face is presented, and finally a blank screen until an overt response is detected or 800 ms have passed. Participants are required to press the left button when identifying a male, and the right button when identifying a female. In Supplementary Figure 1B, correct responses are indicated with black dots, error responses with white dots, and the few right-censored observations with crosses.

@@

Studying the effect of stimulus-onset-asynchrony (SOA) on mean RT is still a popular approach to study the time course of cognitive processes. For example, a target stimulus can be presented with a certain SOA in relation to another stimulus (e.g., prime, mask, distractor, cue). However, mean RT conceals the possibe presence of responses triggered by the non-target stimulus (Panis, 2020; Panis & Schmidt, 2016). For example, in Supplementary Figure 1C a prime is presented before a target with a SOA of -400 ms (upper x-axis labels). To capture responses triggered by the prime (and emitted before target onset) we have to start the timer at trial onset. Also, when RTs are originally recorded relative to target onset, we have to add a constant time to each (here: + 400 ms) in order to include these prime-triggered "negative" RTs in the analysis (lower x-axis labels). In general, all RTs must be positive for EHA to work.

@@

While psychological RT data is typically measured in small, continuous units (e.g., milliseconds), discrete-time EHA treats the RT data as interval-censored data, because it only uses the information that the response occurred sometime in a particular bin of time (x,y]: x < RT $\leq$ y. If we want to use the exact event times, then we treat time as a continuous variable, and let RT be a continous random variable denoting a particular person's response time in a particular trial. Continuous-time EHA does not focus on the cumulative distribution function F(t) = P(RT $\leq$ t) and its derivative, the probability density function f(t) = F(t)', but on the survivor function S(t) = P(RT $>$ t) and the hazard rate function $\lambda$(t) = f(t)/S(t). The hazard rate function gives you the instantaneous *rate* of event occurrence at time point t, given that the event has not occurred yet. Models for continous-time event data come in three forms: parametric models, Cox regression, and piece-wise exponential models [@singerAppliedLongitudinalData2003].

@@

Third, EHA does not discard right-censored observations when estimating hazard functions, that is, trials for which we do not observe a response during the observation period in a trial so that we only know that the RT must be larger than some value (e.g., the response deadline). This is important because although a few right-censored observations are inevitable in most RT tasks, a lot of right-censored observations are expected in experiments on masking, the attentional blink, and so forth. In other words, by using EHA you can analyze RT data from experiments that typically do not measure response times. Ignoring trials without a response leads to underestimation of the true mean. By introducing a fixed censoring time for all trials, trials with long RTs are not discarded but contribute to the risk set of each bin.

### 2025-07-08 manuscript/Suppl_material/supplementary.tex
adjust Suppl. Figure 1

In EHA parlance, the single event of interest is a button press response, \emph{time zero} is defined as target display onset, the \emph{observation period} is 500 ms long in each trial or repeated measurement, in each trial the participant is \emph{at risk} for response occurrence as long as the response has not occurred yet, the individual starts in an ``idle'' state in each trial and \emph{transitions} to a ``detected'' state when a response occurs, all trials have a censoring time of 500 ms, and some trials result in observed event times (those with a RT below 500 ms), while other trials result in response times that are right-censored at 500 ms.

@@

\caption{RT tasks and EHA concepts. The timer starts at time 0 (black vertical line) and ends at the censoring time (dotted vertical line). In each trial of each condition of each task, an overt response time (RT) is indicated by a circle. (A) Single-button detection data requires a single-event analysis. (B) Two-button discrimination data requires a competing risks analysis. (C) Competing risk data in combination with an stimulus-onset-asynchrony manipulation might require increasing the observation period in each trial. (D) Two-button bistable perception data requires a recurrent event / multistate analysis.}\label{fig:plot-RT-tasks}
\end{figure}

@@

Studying the effect of stimulus-onset-asynchrony (SOA) on mean RT is still a popular approach to study the time course of cognitive processes. For example, a target stimulus can be presented with a certain SOA in relation to another stimulus (e.g., prime, mask, distractor, cue). However, mean RT conceals the possibe presence of responses triggered by the non-target stimulus (Panis, 2020; Panis \& Schmidt, 2016). For example, in Supplementary Figure 1C a prime is presented before a target with a SOA of -400 ms (upper x-axis labels). To capture responses triggered by the prime (and emitted before target onset) we have to start the timer at trial onset. Also, when RTs are originally recorded relative to target onset, we have to add a constant time to each (here: + 400 ms) in order to include these ``negative'' RTs in the analysis (lower x-axis labels). In general, all RTs must be positive for EHA to work.

@@

While psychological RT data is typically measured in small, continuous units (e.g., milliseconds), discrete-time EHA treats the RT data as interval-censored data, because it only uses the information that the response occurred sometime in a particular bin of time (x,y{]}: x \textless{} RT \(\leq\) y. If we want to use the exact event times, then we treat time as a continuous variable, and let RT be a continous random variable denoting a particular person's response time in a particular experimental condition. Continuous-time EHA does not focus on the cumulative distribution function F(t) = P(RT \(\leq\) t) and its derivative, the probability density function f(t) = F(t)', but on the survivor function S(t) = P(RT \(>\) t) and the hazard rate function \(\lambda\)(t) = f(t)/S(t). The hazard rate function gives you the instantaneous \emph{rate} of event occurrence at time point t, given that the event has not occurred yet. Models for continous-time event data come in three forms: parametric models, Cox regression, and piece-wise exponential models (Singer \& Willett, 2003).

>>>>

In EHA parlance, the single event of interest is a button press response, \emph{time zero} is defined as target display onset, the \emph{observation or follow-up period} is 500 ms long in each trial or repeated measurement, in each trial the participant is \emph{at risk} for response occurrence as long as the response has not occurred yet, the individual starts in an ``idle'' state in each trial and \emph{transitions} to a ``detected'' state when a response occurs, all trials have a censoring time of 500 ms, and some trials result in observed event times (those with a RT below 500 ms), while other trials result in response times that are right-censored at 500 ms.

@@

\caption{RT tasks and EHA concepts. The timer starts at time 0 (black vertical line) and ends at the censoring time (dotted vertical line). In each trial of each condition of each task, an overt response time (RT) is indicated by a circle. (A) Single-button detection data requires a single-event analysis. (B) Two-button discrimination data requires a competing risks analysis. (C) Competing risk data in combination with an stimulus-onset-asynchrony manipulation might require increasing the observation period in each trial. (D) Two-button bistable perception data requires a recurrent event / multistate analysis. Black bars indicate stimlus presentation.}\label{fig:plot-RT-tasks}
\end{figure}

@@

Studying the effect of stimulus-onset-asynchrony (SOA) on mean RT is still a popular approach to study the time course of cognitive processes. For example, a target stimulus can be presented with a certain SOA in relation to another stimulus (e.g., prime, mask, distractor, cue). However, mean RT conceals the possibe presence of responses triggered by the non-target stimulus (Panis, 2020; Panis \& Schmidt, 2016). For example, in Supplementary Figure 1C a prime is presented before a target with a SOA of -400 ms (upper x-axis labels). To capture responses triggered by the prime (and emitted before target onset) we have to start the timer at trial onset. Also, when RTs are originally recorded relative to target onset, we have to add a constant time to each (here: + 400 ms) in order to include these prime-triggered ``negative'' RTs in the analysis (lower x-axis labels). In general, all RTs must be positive for EHA to work.

@@

While psychological RT data is typically measured in small, continuous units (e.g., milliseconds), discrete-time EHA treats the RT data as interval-censored data, because it only uses the information that the response occurred sometime in a particular bin of time (x,y{]}: x \textless{} RT \(\leq\) y. If we want to use the exact event times, then we treat time as a continuous variable, and let RT be a continous random variable denoting a particular person's response time in a particular trial. Continuous-time EHA does not focus on the cumulative distribution function F(t) = P(RT \(\leq\) t) and its derivative, the probability density function f(t) = F(t)', but on the survivor function S(t) = P(RT \(>\) t) and the hazard rate function \(\lambda\)(t) = f(t)/S(t). The hazard rate function gives you the instantaneous \emph{rate} of event occurrence at time point t, given that the event has not occurred yet. Models for continous-time event data come in three forms: parametric models, Cox regression, and piece-wise exponential models (Singer \& Willett, 2003).

### 2025-07-08 manuscript/reviews/CreateFiguresReview.Rmd
updated packages and re-organised folders
changes to response to reviewers
adjust Suppl. Figure 1

# 2. Create Figure on experimental paradigms

@@

@@

p7 <- data_percept_A %>%
  ggplot(aes(x=RT, y =trial)) +
  geom_vline(xintercept=0) +
  geom_vline(xintercept=max_RT,linetype = 3) +
#    geom_rect(mapping=aes(xmin=0, xmax=100, ymin=7, ymax=6.8), color="black", alpha=0.5) +
  geom_segment(data = lines_a,
               aes(x= xstart, xend = xend,
                   y = ystart, yend = yend),
               linewidth = 1/3,
               linetype = 1) +
  geom_point(shape = c(21,21,21,21,21,21,4,21,21,21,21,21,21,4), 
             size =3, 
             fill =c("black","white","black","white","black","white","black", "white","black","white","black","white","black","black")) + # 4 for x
  scale_x_continuous(limits = c(0,max_RT), breaks=c(0,20,40,60,80,100,120)) +
  scale_y_reverse(breaks = c(1,2),limits = c(3,0), labels = c("1","2"))+
  labs(x = "RT (s)",
       y = "Trial",
       subtitle = "Condition A")+
  ggtitle("Bistable perception task - two buttons") +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
             text = element_text(size = 18, face = "bold"), 
        title = element_text(size = 20, face = "bold"),
        plot.title = element_textbox_simple())

p8 <- data_percept_B %>%
  ggplot(aes(x=RT, y =trial)) +
  geom_vline(xintercept=max_RT,linetype = 3) +
  geom_vline(xintercept=0) +
#    geom_rect(mapping=aes(xmin=0, xmax=100, ymin=7, ymax=6.8), color="black", alpha=0.5) +
  geom_segment(data = lines_b,
               aes(x= xstart, xend = xend,
                   y = ystart, yend = yend),
               linewidth = 1/3,
               linetype = 1) +
  geom_point(shape = c(21,21,21,21,21,21,4,21,21,21,21,21,21,4), size =3, fill =c("white","black","white","black","white","black","black","black","white","black","white","black","white","black")) +
  scale_x_continuous(limits = c(0,max_RT),breaks=c(0,20,40,60,80,100,120)) +
  scale_y_reverse(breaks = c(1:2),limits = c(3,0), labels = c("1","2"))+
  labs(x = "RT (s)",
       y = "Trial",
       subtitle = "Condition B") +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        text = element_text(size = 18.01, face = "bold"), 
        title = element_text(size = 20, face = "bold"))

@@

@@

```{r}
ggsave("REVIEW1/Plot_paradigms.png",dpi = 800,width = 16,height = 17)
```

>>>>

# 2. Create Supplementary Figure 1 on experimental paradigms

@@

create bistable perception plot

@@

p7 <- data_percept_A %>%
  ggplot(aes(x=RT, y =trial)) +
  geom_vline(xintercept=0) +
  geom_vline(xintercept=max_RT,linetype = 3) +
  geom_rect(mapping=aes(xmin=0, xmax=120, ymin=2.8, ymax=2.9), color="black", alpha=0.5) +
  geom_segment(data = lines_a,
               aes(x= xstart, xend = xend,
                   y = ystart, yend = yend),
               linewidth = 1/3,
               linetype = 1) +
  geom_point(shape = c(21,21,21,21,21,21,4,21,21,21,21,21,21,4), 
             size =3, 
             fill =c("black","white","black","white","black","white","black", "white","black","white","black","white","black","black")) + # 4 for x
  scale_x_continuous(limits = c(0,max_RT), breaks=c(0,20,40,60,80,100,120)) +
  scale_y_reverse(breaks = c(1,2),limits = c(3,0), labels = c("1","2"))+
  labs(x = "RT (s)",
       y = "Trial",
       subtitle = "Condition A")+
  ggtitle("Bistable perception task - two buttons") +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
             text = element_text(size = 18, face = "bold"), 
        title = element_text(size = 20, face = "bold"),
        plot.title = element_textbox_simple())

p8 <- data_percept_B %>%
  ggplot(aes(x=RT, y =trial)) +
  geom_vline(xintercept=max_RT,linetype = 3) +
  geom_vline(xintercept=0) +
  geom_rect(mapping=aes(xmin=0, xmax=120, ymin=2.8, ymax=2.9), color="black", alpha=0.5) +
  geom_segment(data = lines_b,
               aes(x= xstart, xend = xend,
                   y = ystart, yend = yend),
               linewidth = 1/3,
               linetype = 1) +
  geom_point(shape = c(21,21,21,21,21,21,4,21,21,21,21,21,21,4), size =3, fill =c("white","black","white","black","white","black","black","black","white","black","white","black","white","black")) +
  scale_x_continuous(limits = c(0,max_RT),breaks=c(0,20,40,60,80,100,120)) +
  scale_y_reverse(breaks = c(1:2),limits = c(3,0), labels = c("1","2"))+
  labs(x = "RT (s)",
       y = "Trial",
       subtitle = "Condition B") +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        text = element_text(size = 18.01, face = "bold"), 
        title = element_text(size = 20, face = "bold"))

@@

create categorization + SOA plot

@@

```{r}
ggsave("manuscript/Suppl_material/Plot_paradigms.png",dpi = 800,width = 16,height = 17)
```
